Competing Bandits: \\ The Perils of Exploration under Competition

### 2018-11-04 paper/content/introduction.tex
comment

\gaedit{We ask whether algorithms that are traditionally considered ``better" in the standard multi-armed bandit problem are incentivized to be deployed under competition.} We investigate this issue via extensive numerical experiments in a stylized duopoly model. In our model, two firms compete for users \gaedit{on a single dimension, quality}, and simultaneously learn from them. Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm. Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window. Each firm's objective is to maximize its \gaedit{expected} market share (the fraction of users choosing this firm).

@@

We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a \gaedit{first-mover}. Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``\gadelete{early entry}\gaedit{incumbent}" (the first-mover in temporary monopoly) to permanent duopoly to ``\gadelete{late entry}\gaedit{entrant" (late-arriver in temporary monopoly)}. \gaedit{We find that, even in the permanent duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late arriver in a market.} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy. \gaedit{Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm that perform better in the long run. As a result, we find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.}

>>>>

\gaedit{We ask whether algorithms that are traditionally considered ``better" in the standard multi-armed bandit problem are incentivized to be deployed under competition.} We investigate this issue via extensive numerical experiments in a stylized duopoly model. In our model, two firms compete for users \gaedit{on a single dimension, quality},\swcomment{not sure what this means} and simultaneously learn from them. Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm. Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window. Each firm's objective is to maximize its \gaedit{expected} market share (the fraction of users choosing this firm).

@@

We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a \gaedit{first-mover}. \swcomment{confused about how they correspond to ``competition intensity''}Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``\gadelete{early entry}\gaedit{incumbent}" (the first-mover in temporary monopoly) to permanent duopoly to ``\gadelete{late entry}\gaedit{entrant" (late-arriver in temporary monopoly)}. \gaedit{We find that, even in the permanent duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late arriver in a market.} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy. \gaedit{Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm that perform better in the long run. As a result, we find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.}
